---
title: "Boyu Chen 678 Midterm Project"
author: "Boyu Chen"
date: "12/4/2021"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(arm)
library(gridExtra)
```

```{r}
df<-read.csv("stats.csv",header = TRUE)
goals<-df$goals
offside<-df$total_offside
ontarget<-df$ontarget_scoring_att
interception<-df$interception
pass<-df$total_pass
season<-df$season
tackle<-df$total_tackle
# Calculate the total credit
win<-df$wins
losse<-df$losses
tie<-38-win-losse
totalpoints<-win*3+tie*1+losse*0
```

Introduction:
The era of data is silently influencing and transforming human life in unimaginable ways, and in sports, the collection and analysis of data is making the sport harmonious and controllable. With the development of technology, soccer has also become surrounded by data and has a wealth of tools to interpret this data. While data analysis may not change the outcome of a game, it can have a huge impact on the outcome of a game. For soccer, the one of the important way to make team more competitive is to use data analytics; and for data analytics to be used in soccer is a testament to its importance.

One example is the penalty that Pogba conceded in the match between Manchester United and Wolves. After collecting and analyzing data on Pogba's penalty taking habits, Wolves' data analysis team found that Pogba would kick the penalty more to the left when his team was in a tie or trailing. Because they told Wolves goalkeeper Patricio in advance, then Patricio was guarding the penalty was moving to the left in advance, and indeed pounced on the penalty. In combination, data analysis can really touch the habits and subconscious performance of players, something that is difficult for players to change and adjust, after all, in many cases it is already set in stone.


Data Processing
In my data, there is not a variable for total points, so I extracted the number of draws, wins, and losses and got the total points by the “number of draws * 1 + number of wins * 3 + number of losses * 0”. In addition, because some variables such as total number of tackles and total number of passes are too big, I use log transfermation for all the variables used in order to make model more accurate.



Here is the variable I used in this project.

```{r}
colnames<-c("Goals","Offside","Ontarget","Pass","Tackle","Total Points","Season")
explaination<-c("Number of goal for each team","Number of offsides in single season","Number of shots on target in single season","Total number of passing ball in single season","Total tackle in single season","Total points in single season","Season")
table <- cbind(colnames, explaination)
colnames(table) <- c("Variables", "Explanation")
knitr::kable(table, "pipe")
```




EDA Part

```{r}
par(mfrow = c(3,3))
ggplot(data=df, mapping=aes(x=log(offside), y=log(totalpoints), group=season)) +
  geom_point(aes(color=season))+
  geom_smooth(se = F,aes(color = season), method = "lm")+
  labs(title = "Total Points vs Total log(offside")


```
Figure 1 is the relationship between log(offside) and log(total points) of each team for each season. According to the plot we can see that the at the most early Premium League(season 2006 - 2007 and season 2008-2012) and most recently Premium League (season 2015-2018) the  log(interception) and total points show the positive relationship. And others show the negative relationship.

```{r}
ggplot(data=df, mapping=aes(x=log(tackle), y=log(totalpoints), group=season)) +
  geom_point(aes(color=season))+
  geom_smooth(se = F,aes(color = season), method = "lm")+
  labs(title = "Total Points vs Total Log(tackles)")
```


Figure 2 is the relationship between log(tackles) and log(total points) for each season.We can see that for almost all the season, higher log(tackles) and number of tackles mean the higher total grades, which means the good defense of team lead to the higher result. However, for season 2017-2018, we can find the negative relationship between log(tackles) and total points. The possible reason that log(tackles) and total points show the negative relationship in season 2017-2018 because the style of the match has some changed


```{r}
ggplot(data=df, mapping=aes(x=log(goals), y=log(totalpoints), group=season)) +
  geom_point(aes(color=season))+
  geom_smooth(se = F,aes(color = season), method = "lm")+
  labs(title = "Total Points vs Total Goals")
```


Figure 3 is the relationship between the number of goals and log(total points) for each season. From the plot we can find that for all the season, the more goals for team leads higher total points. There are obvious positive relationship between goals and total points. The largest slope for the line is season 2008-2009. Because there are an obvious relationship between goals and total points. So I add this indicator to my model.





Model Fitting

Considering different categories, I will use multilevel model to fit the data. All of the variables are continuous variable. And I make log transformation for all the continuous variable and make the log transformation outcomes. Because there is a positive relationship between on target shot and total goals, more goals means the more on target shots. So I add a interaction between log(total goals) and log(on target shot) Below is my function:


```{r}
fit1<-lmer(log(totalpoints)~log(goals)+log(offside)+log(tackle)+log(ontarget)+log(pass)+(1+log(pass)|season)+log(ontarget)*log(goals),data = df)
```


Result

Model Coefficients:

We can get the function here

$$log(totalpoints)= -3.29 + 1.09\cdot log(goals) + 0.1\cdot log(offside) + 0.14\cdot log(tackle) \\
+ 0.07\cdot log(ontarget)+ 0.2\cdot log(pass) + -0.03\cdot log(goals)\cdot log(ontarget)$$

If a team doesn't have any performance on the team, it is obvious that the team will loose the point. 

```{r}
summary(fit1)
# fit2<-lm(totalpoints~goals+interception+offside+ontarget+log(pass) ,data = df)
fit1
plot(fitted(fit1), resid(fit1,type = "response"))
plot(fit1)

AIC(fit1)

re <- plot(fit1)
qq <- lattice::qqmath(fit1)
grid.arrange(re,qq,nrow=1)

coefficients(fit1)
```




Appendix
```{r}
totalgoals<-aggregate(goals, by=list(season=df$season), FUN=sum)
ggplot(data = totalgoals, mapping = aes(x = season, y = x,label = season))+
  geom_bar(stat = 'identity',fill = 'lightblue',width = 0.7)+
  geom_text(mapping = aes(label = x), size = 3, colour = 'red', vjust = 1, hjust = .5, 
            position = position_dodge(0.9))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(x = "season",y = "Total Goals",title = "Total Goals for Each Season")+ylim(0,1100)
```

```{r}
totalpass<-aggregate(pass, by=list(season=df$season), FUN=sum)
ggplot(data = totalpass, mapping = aes(x = season, y = x,label = season))+
  geom_bar(stat = 'identity',fill = 'lightblue',width = 0.7)+
  geom_text(mapping = aes(label = x), size = 3, colour = 'red', vjust = 1, hjust = .5, 
            position = position_dodge(0.9))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(x = "season",y = "Total Pass",title = "Total Pass for Each Season")+ylim(0,380000)
```

```{r}
totalontarget<-aggregate(ontarget, by=list(season=df$season), FUN=sum)
ggplot(data = totalontarget, mapping = aes(x = season, y = x,label = season))+
  geom_bar(stat = 'identity',fill = 'lightblue',width = 0.7)+
  geom_text(mapping = aes(label = x), size = 3, colour = 'red', vjust = 1, hjust = .5, 
            position = position_dodge(0.9))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(x = "season",y = "Total Tackle",title = "Total On Target Shooting For Each Season")+ylim(0,4000)
```

```{r}
totaltackle<-aggregate(tackle, by=list(season=df$season), FUN=sum)
ggplot(data = totaltackle, mapping = aes(x = season, y = x,label = season))+
  geom_bar(stat = 'identity',fill = 'lightblue',width = 0.7)+
  geom_text(mapping = aes(label = x), size = 3, colour = 'red', vjust = 1, hjust = .5, 
            position = position_dodge(0.9))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(x = "season",y = "Total Tackle",title = "Total Tackle For Each Season")+ylim(0,20000)
```

```{r}
totalinterception<-aggregate(interception, by=list(season=df$season), FUN=sum)
ggplot(data = totalinterception, mapping = aes(x = season, y = x,label = season))+
  geom_bar(stat = 'identity',fill = 'lightblue',width = 0.7)+
  geom_text(mapping = aes(label = x), size = 3, colour = 'red', vjust = 1, hjust = .5, 
            position = position_dodge(0.9))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(x = "season",y = "Total Interception",title = "Total Interception For Each Season")+ylim(0,15000)
```

```{r}
totaloffside<-aggregate(offside, by=list(season=df$season), FUN=sum)
ggplot(data = totalinterception, mapping = aes(x = season, y = x,label = season))+
  geom_bar(stat = 'identity',fill = 'lightblue',width = 0.7)+
  geom_text(mapping = aes(label = x), size = 3, colour = 'red', vjust = 1, hjust = .5, 
            position = position_dodge(0.9))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(x = "season",y = "Total Interception",title = "Total Offside For Each Season")+ylim(0,15000)
```

```{r}
ggplot(data=df, mapping=aes(x=goals, y=totalpoints, group=season)) +
 # geom_line(aes(linetype=season,color=season))+
  geom_point(aes(color=season))+
  geom_smooth(se = F,aes(color = season), method = "lm")
```

```{r}
ggplot(data=df, mapping=aes(x=pass, y=totalpoints, group=season)) +
 # geom_line(aes(linetype=season,color=season))+
  geom_point(aes(color=season))+
  geom_smooth(se = F,aes(color = season), method = "lm")
```


```{r}
ggplot(data=df, mapping=aes(x=offside, y=totalpoints, group=season)) +
 # geom_line(aes(linetype=season,color=season))+
  geom_point(aes(color=season))+
  geom_smooth(se = F,aes(color = season), method = "lm")
```


```{r}
ggplot(data=df, mapping=aes(x=ontarget, y=totalpoints, group=season)) +
 # geom_line(aes(linetype=season,color=season))+
  geom_point(aes(color=season))+
  geom_smooth(se = F,aes(color = season), method = "lm")
```

